### 2019-07-29 paper/data/plotsWithTikz.tex
first try on standard deviation

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{pgfplots}
\pgfplotsset{compat=1.8}
\usepgfplotslibrary{statistics}
\usepgfplotslibrary{fillbetween}
\usetikzlibrary{patterns}
\usepackage{amssymb}
\usepackage{subfig}

%\pgfplotscreateplotcyclelist{mylist}{%
%{blue!80!black,mark options={mark size=1.5pt,fill=white},mark=*},
%{red!80!black,densely dashdotted, mark options={mark size=1.5pt,solid,fill=white},mark=square*},
%{green!50!black,mark options={mark size=1.5pt,solid,fill=white},densely dashed,mark=*},
%{yellow!40!black,densely dashdotdotted,mark options={solid,fill=white},mark=triangle*},
%{densely dotted,mark options={solid,fill=white},mark=diamond*}
%}
\tikzstyle{densely dashed}=[dash pattern=on 0.75pt off 0.5pt]

%\definecolor{dark-blue}{rgb}{0,0,0.545}
%\definecolor{dark-red}{rgb}{0.545,0,0}
%\definecolor{dark-green}{rgb}{0,0.392,0}
%\definecolor{dark-magenta}{rgb}{0.545,0,0.545}
%\pgfplotsset{ytick style={draw=o}}
%\renewcommand{\axisdefaultwidth}{300pt}
\renewcommand{\axisdefaultheight}{185pt}

\pgfplotscreateplotcyclelist{mylist}{%
{black,thick,smooth},
{black!50!white,thick,smooth},
{black,thick,densely dashed,smooth},
{black!50!white,thick,densely dashed,smooth},
{black,double,smooth}
}

\pgfplotscreateplotcyclelist{mylist-filled}{%
{black,thick,smooth},
{black,smooth},
{black,smooth},
{black!50!white,thick,smooth},
{black!50!white,smooth},
{black!50!white,smooth},
{black,thick,densely dashed,smooth},
{black,densely dashed,smooth},
{black,densely dashed,smooth},
{black!50!white,thick,densely dashed,smooth},
{black!50!white,densely dashed,smooth},
{black!50!white,densely dashed,smooth},
{black,double,smooth},
{black,double,thin,smooth}
{black,double,thin,smooth}
}

\pgfplotscreateplotcyclelist{fourlist}{%
{black,thick,smooth,mark=x},
{black!50!white,thick,smooth,mark=x},
{black,thick,densely dashed,smooth,mark=x},
{black!50!white,thick,densely dashed,smooth,mark=x}
}

\pgfplotsset{
  /pgfplots/xlabel near ticks/.style={
     /pgfplots/every axis x label/.style={
        at={(ticklabel cs:0.5)},anchor=near ticklabel
     }
  },
  /pgfplots/ylabel near ticks/.style={
     /pgfplots/every axis y label/.style={
        at={(ticklabel cs:0.5)},rotate=90,anchor=near ticklabel}
     }
  }

\begin{document}
\pgfkeys{/pgf/number format/.cd,1000 sep={}}
\begin{figure}
\centering
\subfloat[\label{fig:crossingResolution}{}]{
\centering
\scalebox{.95}{
\centering
%======================
% Crossing Resolution
%======================
\begin{tikzpicture}
\begin{axis}[
			 width=\textwidth,
			 boxplot/draw direction=y,
			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
			 boxplot/box extend=0.18,
			 cycle list name=fourlist,
			 ytick={0,10,20,30,40,50,60,70,80,90},
			 ymajorgrids,
			 xmajorgrids,
			 xtick={0,1,2,...,9},
			 x tick label as interval,
     	     xticklabels={11--20,21--30,31--40,41--50,51--60,
     	     61--70,71--80,81--90,91--100},
     	     ylabel={Crossing Resolution},
			 xlabel={Number of Vertices},
			 ylabel style={yshift=-0.2cm}]
\addplot+[boxplot] table [y=cr-rm, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr11-20.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr21-30.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr21-30.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr31-40.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr31-40.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr41-50.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr41-50.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr51-60.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr51-60.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr61-70.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr61-70.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr71-80.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr71-80.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr81-90.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr81-90.csv};

\addplot+[boxplot] table [y=cr-rm, col sep=semicolon]  {cr91-100.csv};
\addplot+[boxplot] table [y=cr-rm-1, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=cr-fa, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=Crossing Only Crossing Res, col sep=semicolon]{abs-cr91-100.csv};

\end{axis}
\end{tikzpicture}}}

\subfloat[\label{fig:crossingResolution-aspectRatio}{}]{
\centering
\scalebox{.95}{
\centering
%===================================
% Crossing Resolution Aspect Ratio
%===================================
\begin{tikzpicture}
\begin{axis}[
			 width=\textwidth,
			 boxplot/draw direction=y,
			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
			 boxplot/box extend=0.18,
			 cycle list name=fourlist,
			 ytick={1,2.5,10,25,100,250,1000,2500,10000,25000,100000,250000},
			 ymode=log, log ticks with fixed point,
			 ymajorgrids,
			 xmajorgrids,
			 xtick={0,1,2,...,9},
			 x tick label as interval,
     	     xticklabels={11--20,21--30,31--40,41--50,51--60,
     	     61--70,71--80,81--90,91--100},
     	     ylabel style={yshift=-0.2cm},
			 ylabel={Aspect Ratio},
			 xlabel={Number of Vertices}]
\addplot+[boxplot] table [y=asp-rm, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr11-20.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr21-30.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr21-30.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr31-40.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr31-40.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr41-50.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr41-50.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr51-60.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr51-60.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr61-70.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr61-70.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr71-80.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr71-80.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr81-90.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr81-90.csv};

\addplot+[boxplot] table [y=asp-rm, col sep=semicolon]  {cr91-100.csv};
\addplot+[boxplot] table [y=asp-rm-1, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=asp-fa, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=Aspect Ratio, col sep=semicolon]{abs-cr91-100.csv};

\end{axis}
\end{tikzpicture}}}

\end{figure}

\begin{figure}
\centering

\subfloat[\label{fig:crossingResolution-crossings}{}]{
\centering
\scalebox{.95}{
\centering
%======================================
% Crossing Resolution Crossing Number
%======================================
\begin{tikzpicture}
\begin{axis}[
			 width=\textwidth,
			 boxplot/draw direction=y,
     		 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + (1/5)*mod(\plotnumofactualtype,4)},
%			 boxplot/draw position={1/5*(\plotnumofactualtype+1) + 1/5*floor(\plotnumofactualtype/4)},
			 boxplot/box extend=0.18,
			 cycle list name=fourlist,
			 ytick={0,50,100,150,200,250,300,350,400,450,500,550,600,650,700,750,800,850,900,950,1000},
			 ymajorgrids,
			 xmajorgrids,
			 xtick={0,1,2,3,4,5,6,7,8,9},
			 x tick label as interval,
     	     xticklabels={11--20,21--30,31--40,41--50,51--60,
     	     61--70,71--80,81--90,91--100},
     	     ylabel={Number of Crossings},
			 xlabel={Number of Vertices},
			 ylabel style={yshift=-0.2cm},
			]
\addplot+[boxplot] table [y=cn-rm, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr11-20.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr21-30.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr21-30.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr31-40.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr31-40.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr41-50.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr41-50.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr51-60.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr51-60.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr61-70.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr61-70.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr71-80.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr71-80.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr81-90.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr81-90.csv};

\addplot+[boxplot] table [y=cn-rm, col sep=semicolon]  {cr91-100.csv};
\addplot+[boxplot] table [y=cn-rm-1, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=cn-fa, col sep=semicolon] {cr91-100.csv};
\addplot+[boxplot] table [y=Crossing Only No of Crossings, col sep=semicolon]{abs-cr91-100.csv};

\end{axis}
\end{tikzpicture}}}

\subfloat[\label{fig:crossingResolution-iterations}{}]{
\centering
\scalebox{.95}{
\centering
%==================================
% Crossing Resolution Iterations
%==================================
\begin{tikzpicture}
\begin{axis}[
			 width=\textwidth,
			 boxplot/draw direction=y,
			 boxplot/draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*mod(\plotnumofactualtype,4)},
			 boxplot/box extend=0.18,
			 cycle list name=fourlist,
			 ytick={0,5000,10000,15000,20000,25000,30000,35000,40000,45000,50000,55000,  60000,65000},
			 ymajorgrids,
			 xmajorgrids,
			 xtick={0,1,2,...,9},
			 x tick label as interval,
     	    xticklabels={11--20,21--30,31--40,41--50,51--60,
     	     61--70,71--80,81--90,91--100},
     	     ylabel={Iterations},
     	     ylabel style={yshift=-0.2cm},
			 xlabel={Number of Vertices}]
\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr11-20.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr11-20.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr21-30.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr21-30.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr21-30.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr31-40.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr31-40.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr31-40.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr41-50.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr41-50.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr41-50.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr51-60.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr51-60.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr51-60.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr61-70.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr61-70.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr61-70.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr71-80.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr71-80.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr71-80.csv};

\addplot+[boxplot] table [y=iterations-rm, col sep=semicolon]  {cr81-90.csv};
\addplot+[boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=iterations-fa, col sep=semicolon] {cr81-90.csv};
\addplot+[boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr81-90.csv};

\addplot+[mark = x, boxplot] table [y=iterations-rm, col sep=semicolon]  {cr91-100.csv};
\addplot+[mark = x, boxplot] table [y=iterations-rm-1, col sep=semicolon] {cr91-100.csv};
\addplot+[mark = x, boxplot] table [y=iterations-fa, col sep=semicolon] {cr91-100.csv};
\addplot+[mark = x, boxplot] table [y=Crossing Only Iterations, col sep=semicolon]{abs-cr91-100.csv};

\end{axis}
\end{tikzpicture}}}

\end{figure}

\begin{figure*}[t]
\centering
\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
\centering
%======================
% Crossing Resolution
%======================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]

\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};

\addplot[name path=rm_u,black,smooth] table [x=n, y=cr_up rm, col sep=semicolon]{crossingResolution.csv};
\addplot[name path=rm_d,black,smooth] table [x=n, y=cr_down rm, col sep=semicolon]{crossingResolution.csv};
\addplot[black!50,fill opacity=0.5] fill between[of=rm_u and rm_d];

\addplot[name path=rm-1_u,black!50!white,smooth] table [x=n, y=cr_up rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[name path=rm-1_d,black!50!white,smooth] table [x=n, y=cr_down rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[black!25,fill opacity=0.5] fill between[of=rm-1_u and rm-1_d];

\addplot[name path=fa_u,black,smooth,densely dashed] table [x=n, y=cr_up fa, col sep=semicolon]{crossingResolution.csv};
\addplot[name path=fa_d,black,smooth,densely dashed] table [x=n, y=cr_down fa, col sep=semicolon]{crossingResolution.csv};
\addplot[black!50,fill opacity=0.5] fill between[of=fa_u and fa_d];

\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}
\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
\centering
%===================================
% Crossing Resolution Aspect Ratio
%===================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist-filled, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]
\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};
\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}

\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
\centering
%======================================
% Crossing Resolution Crossing Number
%======================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist-filled,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,25,50,75,100,125,150,175,200},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]
\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};
\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}
\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
\centering
%==================================
% Crossing Resolution Iterations
%==================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist-filled,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
xmax=100, ytick={0,500,1000,1500,2000,2500,3000,3500,4000,4500,5000},tick pos=left, ymajorgrids]
\addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};
\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}
\caption{Experimental results on the crossing resolution for the Rome graphs.}
\label{fig:cr-res}
\end{figure*}

\end{document}