### 2019-07-29 paper/data/plotsWithTikz.tex
finished first set of figures

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{pgfplots}
\pgfplotsset{compat=1.8}
\usepgfplotslibrary{statistics}
\usepgfplotslibrary{fillbetween}
\usetikzlibrary{patterns}
\usepackage{amssymb}
\usepackage{subfig}

%\pgfplotscreateplotcyclelist{mylist}{%
%{blue!80!black,mark options={mark size=1.5pt,fill=white},mark=*},
%{red!80!black,densely dashdotted, mark options={mark size=1.5pt,solid,fill=white},mark=square*},
%{green!50!black,mark options={mark size=1.5pt,solid,fill=white},densely dashed,mark=*},
%{yellow!40!black,densely dashdotdotted,mark options={solid,fill=white},mark=triangle*},
%{densely dotted,mark options={solid,fill=white},mark=diamond*}
%}
\tikzstyle{densely dashed}=[dash pattern=on 0.75pt off 0.5pt]

%\definecolor{dark-blue}{rgb}{0,0,0.545}
%\definecolor{dark-red}{rgb}{0.545,0,0}
%\definecolor{dark-green}{rgb}{0,0.392,0}
%\definecolor{dark-magenta}{rgb}{0.545,0,0.545}
%\pgfplotsset{ytick style={draw=o}}
%\renewcommand{\axisdefaultwidth}{300pt}
\renewcommand{\axisdefaultheight}{185pt}

\pgfplotscreateplotcyclelist{mylist}{%
{blue,thick,smooth},
{red,thick,smooth},
{green!70!black,thick,smooth},
{red!40!blue,thick,smooth},
{black,double,smooth}
}

\pgfplotscreateplotcyclelist{mylist-filled}{%
{black,thick,smooth},
{black,smooth},
{black,smooth},
{black!50!white,thick,smooth},
{black!50!white,smooth},
{black!50!white,smooth},
{black,thick,densely dashed,smooth},
{black,densely dashed,smooth},
{black,densely dashed,smooth},
{black!50!white,thick,densely dashed,smooth},
{black!50!white,densely dashed,smooth},
{black!50!white,densely dashed,smooth},
{black,double,smooth},
{black,double,thin,smooth}
{black,double,thin,smooth}
}

\pgfplotscreateplotcyclelist{fourlist}{%
{black,thick,smooth,mark=x},
{black!50!white,thick,smooth,mark=x},
{black,thick,densely dashed,smooth,mark=x},
{black!50!white,thick,densely dashed,smooth,mark=x}
}

\pgfplotsset{
  /pgfplots/xlabel near ticks/.style={
     /pgfplots/every axis x label/.style={
        at={(ticklabel cs:0.5)},anchor=near ticklabel
     }
  },
  /pgfplots/ylabel near ticks/.style={
     /pgfplots/every axis y label/.style={
        at={(ticklabel cs:0.5)},rotate=90,anchor=near ticklabel}
     }
  }

\begin{document}
\pgfkeys{/pgf/number format/.cd,1000 sep={}}

\begin{figure*}[t]
\centering
\subfloat[\label{fig:cr-res-1}{Crossing resolution vs no.~of vertices}]{
\centering
%======================
% Crossing Resolution
%======================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm}, legend style={at={(0.425,1.35)},anchor=north,legend columns=2,draw=none}, width=0.475\textwidth, cycle list name=mylist,
mark repeat={5}, xtick={10,20,30,40,50,60,70,80,90,100}, ytick={0,10,20,30,40,50,60,70,80,90}, ylabel={Crossing Resolution}, xlabel={Number of Vertices}, xmin=10, xmax=100, tick pos=left,ymajorgrids]

\addplot table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};

\addplot[name path=rm_u,blue,smooth,opacity=0.175] table [x=n, y=cr_up rm, col sep=semicolon]{crossingResolution.csv};
\addplot[name path=rm_d,blue,smooth,opacity=0.175] table [x=n, y=cr_down rm, col sep=semicolon]{crossingResolution.csv};
\addplot[blue!80,fill opacity=0.1] fill between[of=rm_u and rm_d];

\addplot[name path=rm-1_u,red!80,smooth,opacity=0.175] table [x=n, y=cr_up rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[name path=rm-1_d,red!80,smooth,opacity=0.175] table [x=n, y=cr_down rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[red!80,fill opacity=0.1] fill between[of=rm-1_u and rm-1_d];

\addplot[name path=fa_u,green!70!black,smooth,opacity=0.175] table [x=n, y=cr_up fa, col sep=semicolon]{crossingResolution.csv};
\addplot[name path=fa_d,,green!70!black,smooth,opacity=0.175] table [x=n, y=cr_down fa, col sep=semicolon]{crossingResolution.csv};
\addplot[green!70!black,fill opacity=0.1] fill between[of=fa_u and fa_d];

\addplot[name path=abs_u,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_up, col sep=semicolon]{abs-crossingResolution.csv};
\addplot[name path=abs_d,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_down, col sep=semicolon]{abs-crossingResolution.csv};
\addplot[red!40!blue,fill opacity=0.1] fill between[of=abs_u and abs_d];

\addplot[blue,smooth,thick] table [x=n, y=crossing resolution rm, col sep=semicolon]{crossingResolution.csv};
\addplot[red,smooth,thick] table [x=n, y=crossing resolution rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[green!70!black,smooth,thick] table [x=n, y=crossing resolution fa, col sep=semicolon]{crossingResolution.csv};
\addplot[red!40!blue,smooth,thick] table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossingResolution.csv};

\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}
\subfloat[\label{fig:cr-res-2}{Aspect ratio vs no.~of vertices}]{
\centering
%===================================
% Crossing Resolution Aspect Ratio
%===================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100},ylabel={Aspect Ratio}, xlabel={Number of Vertices}, ytick={1,2.5,10,25,100,250,1000,2500,10000}, xmin=10, xmax=100,ymode=log, log ticks with fixed point,tick pos=left, ymajorgrids,]

\addplot table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};

\addplot[name path=rm_u,blue,smooth,opacity=0.175] table [x=n, y=asp_up rm, col sep=semicolon]{crossingResolution.csv};
\addplot[name path=rm_d,blue,smooth,opacity=0.175] table [x=n, y=asp_down rm, col sep=semicolon]{crossingResolution.csv};
\addplot[blue!80,fill opacity=0.1] fill between[of=rm_u and rm_d];

\addplot[name path=rm-1_u,red!80,smooth,opacity=0.175] table [x=n, y=asp_up rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[name path=rm-1_d,red!80,smooth,opacity=0.175] table [x=n, y=asp_down rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[red!80,fill opacity=0.1] fill between[of=rm-1_u and rm-1_d];

\addplot[name path=fa_u,green!70!black,smooth,opacity=0.175] table [x=n, y=asp_up fa, col sep=semicolon]{crossingResolution.csv};
\addplot[name path=fa_d,,green!70!black,smooth,opacity=0.175] table [x=n, y=asp_down fa, col sep=semicolon]{crossingResolution.csv};
\addplot[green!70!black,fill opacity=0.1] fill between[of=fa_u and fa_d];

\addplot[name path=abs_u,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_up, col sep=semicolon]{abs-aspectRatio.csv};
\addplot[name path=abs_d,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_down, col sep=semicolon]{abs-aspectRatio.csv};
\addplot[red!40!blue,fill opacity=0.1] fill between[of=abs_u and abs_d];

\addplot[blue,smooth,thick] table [x=n, y=Aspect ratio rm, col sep=semicolon]{crossingResolution.csv};
\addplot[red,smooth,thick] table [x=n, y=Aspect ratio rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[green!70!black,smooth,thick] table [x=n, y=Aspect ratio fa, col sep=semicolon]{crossingResolution.csv};
\addplot[red!40!blue,smooth,thick] table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-aspectRatio.csv};

\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}

\subfloat[\label{fig:cr-res-3}{No.~of crossings vs no.~of vertices}]{
\centering
%======================================
% Crossing Resolution Crossing Number
%======================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10, xmax=100, ytick={0,50,...,350},ylabel={Number of Crossings}, xlabel={Number of Vertices},tick pos=left, ymajorgrids]

\addplot table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};

\addplot[name path=rm_u,blue,smooth,opacity=0.175] table [x=n, y=cn_up rm, col sep=semicolon]{crossingResolution.csv};
\addplot[name path=rm_d,blue,smooth,opacity=0.175] table [x=n, y=cn_down rm, col sep=semicolon]{crossingResolution.csv};
\addplot[blue!80,fill opacity=0.1] fill between[of=rm_u and rm_d];

\addplot[name path=rm-1_u,red!80,smooth,opacity=0.175] table [x=n, y=cn_up rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[name path=rm-1_d,red!80,smooth,opacity=0.175] table [x=n, y=cn_down rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[red!80,fill opacity=0.1] fill between[of=rm-1_u and rm-1_d];

\addplot[name path=fa_u,green!70!black,smooth,opacity=0.175] table [x=n, y=cn_up fa, col sep=semicolon]{crossingResolution.csv};
\addplot[name path=fa_d,,green!70!black,smooth,opacity=0.175] table [x=n, y=cn_down fa, col sep=semicolon]{crossingResolution.csv};
\addplot[green!70!black,fill opacity=0.1] fill between[of=fa_u and fa_d];

\addplot[name path=abs_u,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_up, col sep=semicolon]{abs-crossings.csv};
\addplot[name path=abs_d,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_down, col sep=semicolon]{abs-crossings.csv};
\addplot[red!40!blue,fill opacity=0.1] fill between[of=abs_u and abs_d];

\addplot[blue,smooth,thick] table [x=n, y=Crossing number rm, col sep=semicolon]{crossingResolution.csv};
\addplot[red,smooth,thick] table [x=n, y=Crossing number rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[green!70!black,smooth,thick] table [x=n, y=Crossing number fa, col sep=semicolon]{crossingResolution.csv};
\addplot[red!40!blue,smooth,thick] table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-crossings.csv};

\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}
\subfloat[\label{fig:cr-res-4}{No.~of iterations vs no.~of vertices}]{
\centering
%==================================
% Crossing Resolution Iterations
%==================================
\begin{tikzpicture}
\begin{axis}[axis x line*=bottom,xlabel style={yshift=0.2cm},axis y line*=left,ylabel style={yshift=-0.2cm},legend style={at={(0.425,1.35)}, anchor=north,legend columns=2,draw=none}, width=0.475\textwidth,cycle list name=mylist,  ylabel={Iterations}, xlabel={Number of Vertices}, mark repeat={5},xtick={10,20,30,40,50,60,70,80,90,100}, xmin=10,
xmax=100, ytick={0,2000,4000,...,14000},tick pos=left, ymajorgrids]

\addplot table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
\addplot table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};

\addplot[name path=rm_u,blue,smooth,opacity=0.175] table [x=n, y=it_up rm, col sep=semicolon]{crossingResolution.csv};
\addplot[name path=rm_d,blue,smooth,opacity=0.175] table [x=n, y=it_down rm, col sep=semicolon]{crossingResolution.csv};
\addplot[blue!80,fill opacity=0.1] fill between[of=rm_u and rm_d];

\addplot[name path=rm-1_u,red!80,smooth,opacity=0.175] table [x=n, y=it_up rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[name path=rm-1_d,red!80,smooth,opacity=0.175] table [x=n, y=it_down rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[red!80,fill opacity=0.1] fill between[of=rm-1_u and rm-1_d];

\addplot[name path=fa_u,green!70!black,smooth,opacity=0.175] table [x=n, y=it_up fa, col sep=semicolon]{crossingResolution.csv};
\addplot[name path=fa_d,,green!70!black,smooth,opacity=0.175] table [x=n, y=it_down fa, col sep=semicolon]{crossingResolution.csv};
\addplot[green!70!black,fill opacity=0.1] fill between[of=fa_u and fa_d];

\addplot[name path=abs_u,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_up, col sep=semicolon]{abs-iterations.csv};
\addplot[name path=abs_d,red!40!blue,smooth,opacity=0.175] table [x=Nodes, y=cr_down, col sep=semicolon]{abs-iterations.csv};
\addplot[red!40!blue,fill opacity=0.1] fill between[of=abs_u and abs_d];

\addplot[blue,smooth,thick] table [x=n, y=Iterations rm, col sep=semicolon]{crossingResolution.csv};
\addplot[red,smooth,thick] table [x=n, y=Iterations rm-1, col sep=semicolon] {crossingResolution.csv};
\addplot[green!70!black,smooth,thick] table [x=n, y=Iterations fa, col sep=semicolon]{crossingResolution.csv};
\addplot[red!40!blue,smooth,thick] table [x=Nodes, y=Crossing Only, col sep=semicolon]{abs-iterations.csv};

\legend{{Unrestricted},{AR-restricted~~},{Huang et al.},{Argyriou et al.}}
\end{axis}
\end{tikzpicture}}
\caption{Experimental results on the crossing resolution for the Rome graphs.}
\label{fig:cr-res}
\end{figure*}

\end{document}